\begin{thebibliography}{10}

% \bibitem{Acuna2009}
% V.~Acu{\~{n}}a, F.~Chierichetti, V.~Lacroix, A.~Marchetti-Spaccamela, M.~Sagot, L.~Stougie.
% \newblock {Modes and cuts in metabolic networks: Complexity and algorithms}.
% \newblock {\em Biosystems}, 95(1):51--60, 2009.

\bibitem{anbole13a}
C.~Ans{\'o}tegui, M.~Bonet, J.~Levy.
\newblock {SAT}-based {MaxSAT} algorithms.
\newblock {\em Artificial Intelligence}, 196:77--105, 2013.

\bibitem{baral02a}
C.~Baral.
\newblock {\em Knowledge Representation, Reasoning and Declarative Problem Solving}.
\newblock Cambridge, 2003.

\bibitem{Becker2007}
S.~Becker, A.~Feist, M.~Mo, G.~Hannum, B.~Palsson, M.~Herrgard.
\newblock {Quantitative Prediction of Cellular Metabolism with Constraint-based Models: The COBRA Toolbox}.
\newblock {\em Nature Protocols}, 2(3):727--738, 2007.

\bibitem{coevgeprscsith13a}
G.~Collet, D.~Eveillard, M.~Gebser, S.~Prigent, T.~Schaub, A.~Siegel, S.~Thiele.
\newblock Extending the metabolic network of {E}ctocarpus siliculosus using answer set programming.
\newblock 
% In P.~Cabalar and T.~Son, editors, {\em Proceedings of the Twelfth International Conference on Logic Programming and Nonmonotonic Reasoning (LPNMR'13)}, volume 8148 of {\em Lecture Notes in Artificial Intelligence}, 
{\em Proceedings LPNMR},
245--256. Springer, 2013.

\bibitem{dantzig63a}
G.~Dantzig.
\newblock {\em Linear Programming and Extensions}.
\newblock Princeton, 1963.

% \bibitem{ebhahe04a}
% O.~Ebenhöh, T.~Handorf, R.~Heinrich.
% \newblock Structural analysis of expanding metabolic networks.
% \newblock {\em Genome Informatics}, 15(1):35--45, 2004.

\bibitem{Ebrahim2013}
A.~Ebrahim, J.~Lerman, B.~Palsson, D.~Hyduke.
\newblock {COBRApy: COnstraints-Based Reconstruction and Analysis for Python.}
\newblock {\em BMC Systems Biology}, 7:74, aug 2013.

\bibitem{gekakaosscwa16a}
M.~Gebser, R.~Kaminski, B.~Kaufmann, M.~Ostrowski, T.~Schaub, P.~Wanko.
\newblock Theory solving made easy with clingo~5.
\newblock 
% In M.~Carro and A.~King, editors, {\em Technical Communications of the Thirty-second International Conference on Logic Programming (ICLP'16)}, volume~52, 
{\em Technical Comm.\ ICLP},
2:1--2:15. 
% Open Access Series in Informatics
OASIcs, 2016.

\bibitem{gekakarosc15a}
M.~Gebser, R.~Kaminski, B.~Kaufmann, J.~Romero, T.~Schaub.
\newblock Progress in clasp series 3.
\newblock 
% In F.~Calimeri, G.~Ianni, and M.~Truszczy{\'n}ski, editors, {\em Proceedings of the Thirteenth International Conference on Logic Programming and Nonmonotonic Reasoning (LPNMR'15)}, volume 9345 of {\em Lecture Notes in Artificial Intelligence}, 
{\em Proceedings LPNMR},
368--383. Springer, 2015.

\bibitem{gellif91a}
M.~Gelfond, V.~Lifschitz.
\newblock Classical negation in logic programs and disjunctive databases.
\newblock {\em New Generation Computing}, 9:365--385, 1991.

\bibitem{haebhe05a}
T.~Handorf, O.~Ebenhöh, R.~Heinrich.
\newblock Expanding metabolic networks: Scopes of compounds, robustness, and evolution.
\newblock {\em J.\ of Molec.\ Evolution}, 61(4):498--512, 2005.

\bibitem{laten2014a}
M.~Latendresse.
\newblock Efficiently gap-filling reaction networks.
\newblock {\em BMC bioinformatics}, 15(1):225, 2014.

\bibitem{marzom16a}
C.~Maranas, A.~Zomorrodi.
\newblock {\em Optimiz.\ methods in metabolic networks}.
\newblock Wiley, 2016.

\bibitem{Orth2010}
J.~Orth, B.~Palsson.
\newblock {Systematizing the generation of missing metabolic knowledge.}
\newblock {\em Biotechnology and bioengineering}, 107(3):403--12, oct 2010.

% \bibitem{orthpa10a}
% J.~Orth, I.~Thiele, B.~Palsson.
% \newblock What is flux balance analysis?
% \newblock {\em Nature biotechnology}, 28(3):245--248, 2010.

\bibitem{ostsch12a}
M.~Ostrowski, T.~Schaub.
\newblock {ASP} modulo {CSP}: The clingcon system.
\newblock {\em Theory and Practice of Logic Programming}, 12(4-5):485--503, 2012.

% \bibitem{potassco}
% Potassco website.
% \newblock http://potassco.org.

\bibitem{prcodideetdaevthcabosito14a}
S.~Prigent, G.~Collet, S.~Dittami, L.~Delage, F.~{Ethis de Corny}, O.~Dameron, D.~Eveillard, S.~Thiele, J.~Cambefort, C.~Boyen, A.~Siegel, T.~Tonon.
\newblock The genome-scale metabolic network of ectocarpus siliculosus (ectogem): a resource to study brown algal physiology and beyond.
\newblock {\em The Plant Journal}, 80(2):367–381, 2014.

\bibitem{Prigent2017}
S.~Prigent, C~.Frioux, S.~Dittami, S.~Thiele, A.~Larhlimi, G.~Collet, F.~Gutknecht, J.~Got, D.~Eveillard, J.~Bourdon, F.~Plewniak, T.~Tonon, A.~Siegel.
\newblock {Meneco, a Topology-Based Gap-Filling Tool Applicable to Degraded Genome-Wide Metabolic Networks}.
\newblock {\em PLOS Computational Biology}, 13(1):e1005276, jan 2017.

\bibitem{Reed2003}
J.~Reed, T.~Vo, C.~Schilling, B.~Palsson.
\newblock {An expanded genome-scale model of Escherichia coli K-12 (iJR904 GSM/GPR).}
\newblock {\em Genome Biology}, 4(9):R54, 2003.

\bibitem{SatishKumar2007}
V.~{Satish Kumar}, M.~Dasika, C.~Maranas.
\newblock {Optimization based automated curation of metabolic reconstructions}.
\newblock {\em BMC Bioinformatics}, 8(1):212, 2007.

\bibitem{schthi09a}
T.~Schaub, S.~Thiele.
\newblock Metabolic network expansion with {ASP}.
\newblock
% In P.~Hill and D.~Warren, editors, {\em Proceedings of the Twenty-fifth International Conference on Logic Programming (ICLP'09)}, volume 5649 of {\em Lecture Notes in Computer Science}, 
{\em Proceedings ICLP},
312--326. Springer, 2009.

\bibitem{siniso02a}
P.~Simons, I.~Niemelä, T.~Soininen.
\newblock Extending and implementing the stable model semantics.
\newblock {\em Artificial Intelligence}, 138(1-2):181--234, 2002.

\bibitem{Thiele2014}
I.~Thiele, N.~Vlassis, R.~Fleming.
\newblock {fastGapFill: efficient gap filling in metabolic networks.}
\newblock {\em Bioinformatics}, 30(17):2529--2531, sep 2014.

\bibitem{Vitkin2012}
E.~Vitkin, T.~Shlomi.
\newblock {MIRAGE: a functional genomics-based approach for metabolic network model reconstruction and its application to cyanobacteria networks.}
\newblock {\em Genome Biology}, 13(11):R111, 2012.

\end{thebibliography}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:
