\begin{thebibliography}{49}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{{#1}}
\providecommand{\urlprefix}{URL }
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{Abdullah et~al(2012)Abdullah, Turabieh, McCollum, and
  McMullan}]{DBLP:journals/heuristics/AbdullahTMM12}
Abdullah S, Turabieh H, McCollum B, McMullan P (2012) A hybrid metaheuristic
  approach to the university course timetabling problem. Journal of Heuristics
  18(1):1--23

\bibitem[{Ach{\'a} and Nieuwenhuis(2012)}]{DBLP:journals/aicom/AchaN12}
Ach{\'a} RA, Nieuwenhuis R (2012) Curriculum-based course timetabling with
  {SAT} and {MaxSAT}. Annals of Operations Research pp 1--21

\bibitem[{Andres et~al(2012)Andres, Kaufmann, Matheis, and
  Schaub}]{ankamasc12a}
Andres B, Kaufmann B, Matheis O, Schaub T (2012) Unsatisfiability-based
  optimization in clasp. In: Dovier A, {Santos Costa} V (eds) Technical
  Communications of the Twenty-eighth International Conference on Logic
  Programming (ICLP'12), Leibniz International Proceedings in Informatics
  (LIPIcs), vol~17, pp 212--221

\bibitem[{Ans{\'o}tegui et~al(2013)Ans{\'o}tegui, Bonet, and Levy}]{anbole13a}
Ans{\'o}tegui C, Bonet M, Levy J (2013) {SAT}-based {MaxSAT} algorithms.
  Artificial Intelligence 196:77--105

\bibitem[{Banbara et~al(2013)Banbara, Soh, Tamura, Inoue, and
  Schaub}]{basotainsc13a}
Banbara M, Soh T, Tamura N, Inoue K, Schaub T (2013) Answer set programming as
  a modeling language for course timetabling. Theory and Practice of Logic
  Programming 13(4-5):783--798

\bibitem[{Banbara et~al(2016)Banbara, Inoue, Kaufmann, Schaub, Soh, Tamura, and
  Wanko}]{bainkascsotawa16}
Banbara M, Inoue K, Kaufmann B, Schaub T, Soh T, Tamura N, Wanko P (2016)
  teaspoon: Solving the curriculum-based course timetabling problems with
  answer set programming. In: Burke EK, {Di Gaspero} L, {\"{O}}zcan E, McCollum
  B, Schaerf A (eds) Proceedings of the 11th International Conference on the
  Practice and Theory of Automated Timetabling (PATAT'16), pp 13--32

\bibitem[{Baral(2003)}]{baral03:cambridge}
Baral C (2003) Knowledge Representation, Reasoning and Declarative Problem
  Solving. Cambridge University Press

\bibitem[{Bart{\'{a}}k et~al(2004)Bart{\'{a}}k, M{\"{u}}ller, and
  Rudov{\'{a}}}]{BartakMR03}
Bart{\'{a}}k R, M{\"{u}}ller T, Rudov{\'{a}} H (2004) A new approach to
  modeling and solving minimal perturbation problems. In: Apt KR, Fages F,
  Rossi F, Szeredi P, V{\'{a}}ncza J (eds) Recent Advances in Constraints,
  Joint ERCIM/CoLogNET International Workshop on Constraint Solving and
  Constraint Logic Programming, (CSCLP'03), Springer, LNCS, vol 3010, pp
  233--249

\bibitem[{Bettinelli et~al(2015)Bettinelli, Cacchiani, Roberti, Paolo, and
  Toth}]{Bettinelli2015}
Bettinelli A, Cacchiani V, Roberti R, Paolo, Toth (2015) An overview of
  curriculum-based course timetabling. TOP 23(2):313--349

\bibitem[{Biere et~al(2009)Biere, Heule, {van Maaren}, and Walsh}]{SATHandbook}
Biere A, Heule M, {van Maaren} H, Walsh T (eds)  (2009) Handbook of
  Satisfiability, Frontiers in Artificial Intelligence and Applications, vol
  185. IOS Press

\bibitem[{Bonutti et~al(2012)Bonutti, {De Cesco}, {Di Gaspero}, and
  Schaerf}]{DBLP:journals/anor/BonuttiCGS12}
Bonutti A, {De Cesco} F, {Di Gaspero} L, Schaerf A (2012) Benchmarking
  curriculum-based course timetabling: formulations, data formats, instances,
  validation, visualization, and results. Annals of Operations Research
  194(1):59--70

\bibitem[{Burke and Petrovic(2002)}]{DBLP:journals/eor/BurkeP02}
Burke EK, Petrovic S (2002) Recent research directions in automated
  timetabling. European Journal of Operational Research 140(2):266--280

\bibitem[{Burke et~al(2010{\natexlab{a}})Burke, Marecek, Parkes, and
  Rudov{\'a}}]{DBLP:journals/cor/BurkeMPR10}
Burke EK, Marecek J, Parkes AJ, Rudov{\'a} H (2010{\natexlab{a}})
  Decomposition, reformulation, and diving in university course timetabling.
  Computers {\&} Operations Research 37(3):582--597

\bibitem[{Burke et~al(2010{\natexlab{b}})Burke, Marecek, Parkes, and
  Rudov{\'a}}]{DBLP:journals/anor/BurkeMPR10}
Burke EK, Marecek J, Parkes AJ, Rudov{\'a} H (2010{\natexlab{b}}) A supernodal
  formulation of vertex colouring with applications in course timetabling.
  Annals of Operations Research 179(1):105--130

\bibitem[{Burke et~al(2012)Burke, Marecek, Parkes, and
  Rudov{\'a}}]{DBLP:journals/anor/BurkeMPR12}
Burke EK, Marecek J, Parkes AJ, Rudov{\'a} H (2012) A branch-and-cut procedure
  for the Udine course timetabling problem. Annals of Operations Research
  194(1):71--87

\bibitem[{Demirovic and Musliu(2017)}]{DBLP:journals/cor/DemirovicM17}
Demirovic E, Musliu N (2017) MaxSAT-based large neighborhood search for high
  school timetabling. Computers {\&} {OR} 78:172--180

\bibitem[{{Di Gaspero} and Schaerf(2003)}]{DBLP:conf/patat/GasperoS02}
{Di Gaspero} L, Schaerf A (2003) Multi-neighbourhood local search with
  application to course timetabling. In: Burke EK, Causmaecker PD (eds)
  Proceedings of the 4th International Conference on the Practice and Theory of
  Automated Timetabling (PATAT'02), Springer, Berlin Heidelberg, 
  LNCS, vol 2740, pp 262--275

\bibitem[{{Di Gaspero} and Schaerf(2006)}]{DBLP:journals/jmma/GasperoS06}
{Di Gaspero} L, Schaerf A (2006) Neighborhood portfolio approach for local
  search applied to timetabling problems. Journal of Mathematical Modelling and
  Algorithms 5(1):65--89

\bibitem[{{Di Gaspero} et~al(2007){Di Gaspero}, McCollum, and
  Schaerf}]{GasperoMS/ITC2007}
{Di Gaspero} L, McCollum B, Schaerf A (2007) The second international
  timetabling competition ({ITC}-2007): Curriculum-based course timetabling
  (track 3). Technical report, Queen's University, Belfast, United Kingdom

\bibitem[{E{\'{e}}n and S{\"{o}}rensson(2003)}]{DBLP:journals/entcs/EenS03}
E{\'{e}}n N, S{\"{o}}rensson N (2003) Temporal induction by incremental {SAT}
  solving. Electronic Notes in Theoretical Computer Science 89(4):543--560

\bibitem[{Ehrgott(2005)}]{Ehrgott2005}
Ehrgott M (2005) Multicriteria Optimization. Springer Berlin Heidelberg

\bibitem[{Erdem et~al(2016)Erdem, Gelfond, and Leone}]{ergele16a}
Erdem E, Gelfond M, Leone N (2016) Applications of {ASP}. AI Magazine
  37(3):53--68

\bibitem[{Faber et~al(1998)Faber, Leone, and Pfeifer}]{Faber98}
Faber W, Leone N, Pfeifer G (1998) Representing school timetabling in a
  disjunctive logic programming language. In: Egly U, Tompits H (eds)
  Proceedings of the 13th Workshop on Logic Programming (WLP'98), pp 43--52

\bibitem[{Gebser et~al(2012)Gebser, Kaminski, Kaufmann, and
  Schaub}]{gekakasc12a}
Gebser M, Kaminski R, Kaufmann B, Schaub T (2012) Answer Set Solving in
  Practice. Synthesis Lectures on Artificial Intelligence and Machine Learning,
  Morgan and Claypool Publishers

\bibitem[{Gebser et~al(2013)Gebser, Kaufmann, Otero, Romero, Schaub, and
  Wanko}]{gekaotroscwa13a}
Gebser M, Kaufmann B, Otero R, Romero J, Schaub T, Wanko P (2013)
  Domain-specific heuristics in answer set programming. In: {desJardins} M,
  Littman M (eds) Proceedings of the Twenty-Seventh National Conference on
  Artificial Intelligence (AAAI'13), AAAI Press, pp 350--356

\bibitem[{Gebser et~al(2015{\natexlab{a}})Gebser, Kaminski, Kaufmann, Romero,
  and Schaub}]{gekakarosc15a}
Gebser M, Kaminski R, Kaufmann B, Romero J, Schaub T (2015{\natexlab{a}})
  Progress in clasp series 3. In: Calimeri F, Ianni G, Truszczy{\'n}ski M (eds)
  Proceedings of the Thirteenth International Conference on Logic Programming
  and Nonmonotonic Reasoning (LPNMR'15), Springer-Verlag, LNAI, vol 9345, pp 368--383

\bibitem[{Gebser et~al(2015{\natexlab{b}})Gebser, Kaminski, Obermeier, and
  Schaub}]{gekaobsc15a}
Gebser M, Kaminski R, Obermeier P, Schaub T (2015{\natexlab{b}}) Ricochet
  robots reloaded: A case-study in multi-shot {ASP} solving. In: Eiter T,
  Strass H, Truszczy{\'n}ski M, Woltran S (eds) Advances in Knowledge
  Representation, Logic Programming, and Abstract Argumentation: Essays
  Dedicated to {G}erhard {B}rewka on the Occasion of His 60th Birthday,
  Springer-Verlag, LNAI, vol 9060, pp 17--32

\bibitem[{Geiger(2012)}]{DBLP:journals/anor/Geiger12}
Geiger MJ (2012) Applying the threshold accepting metaheuristic to curriculum
  based course timetabling - a contribution to the second international
  timetabling competition ITC 2007. Annals of Operations Research
  194(1):189--202

\bibitem[{Gelfond and Lifschitz(1988)}]{DBLP:conf/iclp/GelfondL88}
Gelfond M, Lifschitz V (1988) The stable model semantics for logic programming.
  In: Proceedings of the Fifth International Conference and Symposium on Logic
  Programming, MIT Press, pp 1070--1080

\bibitem[{Hutter et~al(2011)Hutter, Hoos, and Leyton-Brown}]{huhole11b}
Hutter F, Hoos H, Leyton-Brown K (2011) Sequential model-based optimization for
  general algorithm configuration. In: Proceedings of the Fifth International
  Conference on Learning and Intelligent Optimization (LION'11),
  Springer-Verlag, LNCS, vol 6683, pp 507--523

\bibitem[{Lach and L{\"u}bbecke(2012)}]{DBLP:journals/anor/LachL12}
Lach G, L{\"u}bbecke ME (2012) Curriculum based course timetabling: new
  solutions to Udine benchmark instances. Annals of Operations Research
  194(1):255--272

\bibitem[{Lewis(2007)}]{Lewis2007:survey}
Lewis R (2007) A survey of metaheuristic-based techniques for university
  timetabling problems. OR Spectrum 30(1):167--190

\bibitem[{L{\"u} and Hao(2010)}]{DBLP:journals/eor/LuH10}
L{\"u} Z, Hao JK (2010) Adaptive tabu search for course timetabling. European
  Journal of Operational Research 200(1):235--244

\bibitem[{Marler and Arora(2004)}]{Marler2004}
Marler R, Arora J (2004) Survey of multi-objective optimization methods for
  engineering. Structural and Multidisciplinary Optimization 26(6):369--395

\bibitem[{Marques-Silva and Planes(2007)}]{marpla07}
Marques-Silva J, Planes J (2007) On using unsatisfiability for solving maximum
  satisfiability. CoRR abs/0712.1097

\bibitem[{McCollum(2007)}]{DBLP:conf/patat/McCollum06}
McCollum B (2007) A perspective on bridging the gap between theory and practice
  in university timetabling. In: Burke EK, Rudov{\'a} H (eds) Proceedings of
  the 6th International Conference on the Practice and Theory of Automated
  Timetabling (PATAT'06), Revised Selected Papers, Springer, LNCS, vol 3867, pp 3--23

\bibitem[{McCollum et~al(2010)McCollum, Schaerf, Paechter, McMullan, Lewis,
  Parkes, {Di Gaspero}, Qu, and
  Burke}]{DBLP:journals/informs/McCollumSPMLPGQB10}
McCollum B, Schaerf A, Paechter B, McMullan P, Lewis R, Parkes AJ, {Di Gaspero}
  L, Qu R, Burke EK (2010) Setting the research agenda in automated
  timetabling: The second international timetabling competition. INFORMS
  Journal on Computing 22(1):120--130

\bibitem[{M{\"u}ller(2009)}]{DBLP:journals/anor/Muller09}
M{\"u}ller T (2009) {ITC2007} solver description: a hybrid approach. Annals of
  Operations Research 172(1):429--446

\bibitem[{M{\"{u}}ller et~al(2005)M{\"{u}}ller, Rudov{\'{a}}, and
  Bart{\'{a}}k}]{MullerRB04}
M{\"{u}}ller T, Rudov{\'{a}} H, Bart{\'{a}}k R (2005) Minimal perturbation
  problem in course timetabling. In: Burke EK, Trick MA (eds) Proceedings of
  the 5th International Conference on the Practice and Theory of Automated
  Timetabling (PATAT'04), Springer, LNCS, vol 3616, pp 126--146

\bibitem[{Narodytska and Bacchus(2014)}]{narbac14a}
Narodytska N, Bacchus F (2014) Maximum satisfiability using core-guided MaxSAT
  resolution. In: Brodley C, Stone P (eds) Proceedings of the Twenty-Eighth
  National Conference on Artificial Intelligence (AAAI'14), AAAI Press, pp
  2717--2723

\bibitem[{Niemel{\"a}(1999)}]{DBLP:journals/amai/Niemela99}
Niemel{\"a} I (1999) Logic programs with stable model semantics as a constraint
  programming paradigm. Ann\ Mathematics and Artificial Intelligence
  25(3--4):241--273

\bibitem[{Phillips et~al(2016)Phillips, Walker, Ehrgott, and
  Ryan}]{Phillips2016}
Phillips AE, Walker CG, Ehrgott M, Ryan DM (2016) Integer programming for
  minimal perturbation problems in university course timetabling. Annals of
  Operations Research pp 1--22

\bibitem[{Rudov{\'{a}} et~al(2011)Rudov{\'{a}}, M{\"{u}}ller, and
  Murray}]{RudovaMM11}
Rudov{\'{a}} H, M{\"{u}}ller T, Murray KS (2011) Complex university course
  timetabling. Journal of Scheduling 14(2):187--207

\bibitem[{Sakkout and Wallace(2000)}]{Hani2000}
Sakkout HE, Wallace M (2000) Probe backtrack search for minimal perturbation in
  dynamic scheduling. Constraints 4(5):359--388

\bibitem[{Schaerf(1999)}]{DBLP:journals/air/Schaerf99}
Schaerf A (1999) A survey of automated timetabling. Artificial Intelligence
  Review 13(2):87--127

\bibitem[{Schwind et~al(2014)Schwind, Okimoto, Konieczny, Wack, and
  Inoue}]{DBLP:conf/ictai/SchwindOKWI14}
Schwind N, Okimoto T, Konieczny S, Wack M, Inoue K (2014) Utilitarian and
  egalitarian solutions for multi-objective constraint optimization. In:
  Proceedings of the 26th {IEEE} International Conference on Tools with
  Artificial Intelligence (ICTAI'14), {IEEE} Computer Society, pp 170--177

\bibitem[{Sinz(2005)}]{sinz05a}
Sinz C (2005) Towards an optimal {CNF} encoding of {B}oolean cardinality
  constraints. In: {van Beek} P (ed) Proceedings of the Eleventh International
  Conference on Principles and Practice of Constraint Programming (CP'05),
  Springer-Verlag, LNCS, vol 3709, pp 827--831

\bibitem[{Soh et~al(2017)Soh, Banbara, Tamura, and Berre}]{sbtl2017}
Soh T, Banbara M, Tamura N, Berre DL (2017) Solving multiobjective discrete
  optimization problems with propositional minimal model generation. In: Beck
  JC (ed) Proceedings of the 23rd International Conference on Principles and
  Practice of Constraint Programming (CP'17), Springer, LNCS, vol 10416, pp 596--614

\bibitem[{Zivan et~al(2011)Zivan, Grubshtein, and Meisels}]{Roie2011}
Zivan R, Grubshtein A, Meisels A (2011) Hybrid search for minimal perturbation
  in dynamic CSPs. Constraints 16(3):228--249
\end{thebibliography}
