\begin{abstract}
Answer Set Programming (ASP) has become an established paradigm for Knowledge Representation and Reasoning,
in particular, when it comes to solving knowledge-intense combinatorial (optimization) problems.
ASP's unique pairing of a simple yet rich modeling language with highly performant solving technology
has led to an increasing interest in ASP in academia as well as industry.
To further boost this development and make ASP fit for real world applications it is indispensable to equip it with means 
for an easy integration into software environments and for adding complementary forms of reasoning.

In this tutorial, we describe how both issues are addressed in the ASP system \clingo.
At first, we outline features of \clingo's application programming interface (API) that are essential for multi-shot ASP solving,
a technique for dealing with continuously changing logic programs.
This is illustrated by realizing two exemplary reasoning modes, namely branch-and-bound-based optimization and incremental ASP solving.
We then switch to the design of the API for integrating complementary forms of reasoning and detail this in an extensive case study
dealing with the integration of difference constraints.
We show how the syntax of these constraints is added to the modeling language and seamlessly merged into the grounding process.
We then develop in detail a corresponding theory propagator for difference constraints and present how it is integrated into \clingo's solving process.
\end{abstract}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: 
