\begin{thebibliography}{10}

\bibitem{ballie17a}
M.~Balduccini and Y.~Lierler.
\newblock Constraint answer set solver {EZCSP} and why integration schemas
  matter.
\newblock {\em CoRR}, abs/1702.04047, 2017.

\bibitem{bageinospescsotawe15a}
M.~Banbara, M.~Gebser, K.~Inoue, M.~Ostrowski, A.~Peano, T.~Schaub, T.~Soh,
  N.~Tamura, and M.~Weise.
\newblock aspartame: Solving constraint satisfaction problems with answer set
  programming.
\newblock In F.~Calimeri, G.~Ianni, and M.~Truszczy{\'n}ski, editors, {\em
  Proceedings of the Thirteenth International Conference on Logic Programming
  and Nonmonotonic Reasoning (LPNMR'15)}, volume 9345 of {\em Lecture Notes in
  Artificial Intelligence}, pages 112--126. Springer-Verlag, 2015.

\bibitem{bakaossc16a}
M.~Banbara, B.~Kaufmann, M.~Ostrowski, and T.~Schaub.
\newblock Clingcon: The next generation.
\newblock {\em Theory and Practice of Logic Programming}, 2017.
\newblock To appear.

\bibitem{baral02a}
C.~Baral.
\newblock {\em Knowledge Representation, Reasoning and Declarative Problem
  Solving}.
\newblock Cambridge University Press, 2003.

\bibitem{baseseti09a}
C.~Barrett, R.~Sebastiani, S.~Seshia, and C.~Tinelli.
\newblock Satisfiability modulo theories.
\newblock In Biere et~al. \cite{SATHandbook}, chapter~26, pages 825--885.

\bibitem{barlee14b}
M.~Bartholomew and J.~Lee.
\newblock System aspmt2smt: Computing {ASPMT} theories by {SMT} solvers.
\newblock In E.~Ferm\'{e} and J.~Leite, editors, {\em Proceedings of the
  Fourteenth European Conference on Logics in Artificial Intelligence
  (JELIA'14)}, volume 8761 of {\em Lecture Notes in Artificial Intelligence},
  pages 529--542. Springer-Verlag, 2014.

\bibitem{SATHandbook}
A.~Biere, M.~Heule, H.~{van Maaren}, and T.~Walsh, editors.
\newblock {\em Handbook of Satisfiability}, volume 185 of {\em Frontiers in
  Artificial Intelligence and Applications}.
\newblock IOS Press, 2009.

\bibitem{cakaossc16a}
P.~Cabalar, R.~Kaminski, M.~Ostrowski, and T.~Schaub.
\newblock An {ASP} semantics for default reasoning with constraints.
\newblock In R.~Kambhampati, editor, {\em Proceedings of the Twenty-fifth
  International Joint Conference on Artificial Intelligence (IJCAI'16)}, pages
  1015--1021. IJCAI/AAAI Press, 2016.

\bibitem{iclp-lipics16}
M.~Carro and A.~King, editors.
\newblock {\em Technical Communications of the Thirty-second International
  Conference on Logic Programming (ICLP'16)}, volume~52. Open Access Series in
  Informatics (OASIcs), 2016.

\bibitem{cotmal06a}
S.~Cotton and O.~Maler.
\newblock Fast and flexible difference constraint propagation for {DPLL (T)}.
\newblock In A.~Biere and C.~Gomes, editors, {\em Proceedings of the Ninth
  International Conference on Theory and Applications of Satisfiability Testing
  (SAT'06)}, volume 4121 of {\em Lecture Notes in Computer Science}, pages
  170--183. Springer-Verlag, 2006.

\bibitem{crabak94a}
J.~Crawford and A.~Baker.
\newblock Experimental results on the application of satisfiability algorithms
  to scheduling problems.
\newblock In B.~Hayes-Roth and R.~Korf, editors, {\em Proceedings of the
  Twelfth National Conference on Artificial Intelligence (AAAI'94)}, pages
  1092--1097. AAAI Press, 1994.

\bibitem{dorisc16a}
C.~Dodaro, F.~Ricca, and P.~Schüller.
\newblock External propagators in wasp: Preliminary report.
\newblock In {\em Proceedings of the Twenty-third International Workshop on
  Experimental Evaluation of Algorithms for Solving Problems with Combinatorial
  Explosion (RCRA'16)}, volume 1745, pages 1--9. CEUR Workshop Proceedings,
  2016.

\bibitem{drewal10a}
C.~Drescher and T.~Walsh.
\newblock A translational approach to constraint answer set solving.
\newblock {\em Theory and Practice of Logic Programming}, 10(4-6):465--480,
  2010.

\bibitem{eiiakr09a}
T.~Eiter, G.~Ianni, and T.~Krennwallner.
\newblock {A}nswer {S}et {P}rogramming: {A} {P}rimer.
\newblock In S.~Tessaris, E.~Franconi, T.~Eiter, C.~Gutierrez, S.~Handschuh,
  M.~Rousset, and R.~Schmidt, editors, {\em Fifth International Reasoning Web
  Summer School (RW'09)}, volume 5689 of {\em Lecture Notes in Computer
  Science}, pages 40--110. Springer-Verlag, 2009.

\bibitem{gegrkasc11a}
M.~Gebser, T.~Grote, R.~Kaminski, and T.~Schaub.
\newblock Reactive answer set programming.
\newblock In J.~Delgrande and W.~Faber, editors, {\em Proceedings of the
  Eleventh International Conference on Logic Programming and Nonmonotonic
  Reasoning (LPNMR'11)}, volume 6645 of {\em Lecture Notes in Artificial
  Intelligence}, pages 54--66. Springer-Verlag, 2011.

\bibitem{PotasscoUserGuide}
M.~Gebser, R.~Kaminski, B.~Kaufmann, M.~Lindauer, M.~Ostrowski, J.~Romero,
  T.~Schaub, and S.~Thiele.
\newblock {\em Potassco User Guide}.
\newblock University of Potsdam, second edition edition, 2015.

\bibitem{gekakaosscth08a}
M.~Gebser, R.~Kaminski, B.~Kaufmann, M.~Ostrowski, T.~Schaub, and S.~Thiele.
\newblock Engineering an incremental {ASP} solver.
\newblock In M.~{Garcia de la Banda} and E.~Pontelli, editors, {\em Proceedings
  of the Twenty-fourth International Conference on Logic Programming
  (ICLP'08)}, volume 5366 of {\em Lecture Notes in Computer Science}, pages
  190--205. Springer-Verlag, 2008.

\bibitem{gekakaosscwa16a}
M.~Gebser, R.~Kaminski, B.~Kaufmann, M.~Ostrowski, T.~Schaub, and P.~Wanko.
\newblock Theory solving made easy with clingo~5.
\newblock In Carro and King \cite{iclp-lipics16}, pages 2:1--2:15.

\bibitem{gekakasc12a}
M.~Gebser, R.~Kaminski, B.~Kaufmann, and T.~Schaub.
\newblock {\em Answer Set Solving in Practice}.
\newblock Synthesis Lectures on Artificial Intelligence and Machine Learning.
  Morgan and Claypool Publishers, 2012.

\bibitem{gekakasc14b}
M.~Gebser, R.~Kaminski, B.~Kaufmann, and T.~Schaub.
\newblock \textit{Clingo} = {ASP} + control: Preliminary report.
\newblock In M.~Leuschel and T.~Schrijvers, editors, {\em Technical
  Communications of the Thirtieth International Conference on Logic Programming
  (ICLP'14)}, volume arXiv:1405.3694v1 of {\em Theory and Practice of Logic
  Programming, Online Supplement}, 2014.
\newblock Available at \url{http://arxiv.org/abs/1405.3694v1}.

\bibitem{gekaobsc15a}
M.~Gebser, R.~Kaminski, P.~Obermeier, and T.~Schaub.
\newblock Ricochet robots reloaded: A case-study in multi-shot {ASP} solving.
\newblock In T.~Eiter, H.~Strass, M.~Truszczy{\'n}ski, and S.~Woltran, editors,
  {\em Advances in Knowledge Representation, Logic Programming, and Abstract
  Argumentation: Essays Dedicated to {G}erhard {B}rewka on the Occasion of His
  60th Birthday}, volume 9060 of {\em Lecture Notes in Artificial
  Intelligence}, pages 17--32. Springer-Verlag, 2015.

\bibitem{gekasc09c}
M.~Gebser, B.~Kaufmann, and T.~Schaub.
\newblock Conflict-driven answer set solving: From theory to practice.
\newblock {\em Artificial Intelligence}, 187-188:52--89, 2012.

\bibitem{gelkah14a}
M.~Gelfond and Y.~Kahl.
\newblock {\em Knowledge Representation, Reasoning, and the Design of
  Intelligent Agents: The Answer-Set Programming Approach}.
\newblock Cambridge University Press, 2014.

\bibitem{gellif88b}
M.~Gelfond and V.~Lifschitz.
\newblock The stable model semantics for logic programming.
\newblock In R.~Kowalski and K.~Bowen, editors, {\em Proceedings of the Fifth
  International Conference and Symposium of Logic Programming (ICLP'88)}, pages
  1070--1080. MIT Press, 1988.

\bibitem{gellif91a}
M.~Gelfond and V.~Lifschitz.
\newblock Classical negation in logic programs and disjunctive databases.
\newblock {\em New Generation Computing}, 9:365--385, 1991.

\bibitem{jakaosscscwa17a}
T.~Janhunen, R.~Kaminski, M.~Ostrowski, T.~Schaub, S.~Schellhorn, and P.~Wanko.
\newblock Clingo goes linear constraints over reals and integers: A preliminary
  study.
\newblock 2017.
\newblock In preparation.

\bibitem{jalini11a}
T.~Janhunen, G.~Liu, and I.~Niemelä.
\newblock Tight integration of non-ground answer set programming and
  satisfiability modulo theories.
\newblock In P.~Cabalar, D.~Mitchell, D.~Pearce, and E.~Ternovska, editors,
  {\em Proceedings of the First Workshop on Grounding and Transformation for
  Theories with Variables (GTTV'11)}, pages 1--13, 2011.

\bibitem{liman09a}
C.~Li and F.~Many{\`a}.
\newblock Max{SAT}.
\newblock In Biere et~al. \cite{SATHandbook}, chapter~19, pages 613--631.

\bibitem{liesus16a}
Y.~Lierler and B.~Susman.
\newblock {SMT}-based constraint answer set solver {EZSMT} (system
  description).
\newblock In Carro and King \cite{iclp-lipics16}, pages 1:1--1:15.

\bibitem{lifschitz04a}
V.~Lifschitz.
\newblock Introduction to answer set programming.
\newblock Unpublished draft, 2004.

\bibitem{lijani12a}
G.~Liu, T.~Janhunen, and I.~Niemelä.
\newblock Answer set programming via mixed integer programming.
\newblock In G.~Brewka, T.~Eiter, and S.~McIlraith, editors, {\em Proceedings
  of the Thirteenth International Conference on Principles of Knowledge
  Representation and Reasoning (KR'12)}, pages 32--42. AAAI Press, 2012.

\bibitem{malyma09a}
J.~Marques-Silva, I.~Lynce, and S.~Malik.
\newblock Conflict-driven clause learning {SAT} solvers.
\newblock In Biere et~al. \cite{SATHandbook}, chapter~4, pages 131--153.

\bibitem{megezh08a}
V.~Mellarkod, M.~Gelfond, and Y.~Zhang.
\newblock Integrating answer set programming and constraint logic programming.
\newblock {\em Annals of Mathematics and Artificial Intelligence},
  53(1-4):251--287, 2008.

\bibitem{niolti06a}
R.~Nieuwenhuis, A.~Oliveras, and C.~Tinelli.
\newblock Solving {SAT} and {SAT} modulo theories: From an abstract
  {D}avis-{P}utnam-{L}ogemann-{L}oveland procedure to {DPLL}({T}).
\newblock {\em Journal of the ACM}, 53(6):937--977, 2006.

\bibitem{oikjan06a}
E.~Oikarinen and T.~Janhunen.
\newblock Modular equivalence for normal logic programs.
\newblock In G.~Brewka, S.~Coradeschi, A.~Perini, and P.~Traverso, editors,
  {\em Proceedings of the Seventeenth European Conference on Artificial
  Intelligence (ECAI'06)}, pages 412--416. IOS Press, 2006.

\bibitem{redl16a}
C.~Redl.
\newblock The dlvhex system for knowledge representation: recent advances
  (system description).
\newblock {\em Theory and Practice of Logic Programming}, 16(5-6):866--883,
  2016.

\bibitem{rouman09a}
O.~Roussel and V.~Manquinho.
\newblock Pseudo-{B}oolean and cardinality constraints.
\newblock In Biere et~al. \cite{SATHandbook}, chapter~22, pages 695--733.

\bibitem{siniso02a}
P.~Simons, I.~Niemelä, and T.~Soininen.
\newblock Extending and implementing the stable model semantics.
\newblock {\em Artificial Intelligence}, 138(1-2):181--234, 2002.

\bibitem{lparseManual}
T.~Syrjänen.
\newblock Lparse 1.0 user's manual, 2001.

\bibitem{tatakiba09a}
N.~Tamura, A.~Taga, S.~Kitagawa, and M.~Banbara.
\newblock Compiling finite linear {CSP} into {SAT}.
\newblock {\em Constraints}, 14(2):254--272, 2009.

\end{thebibliography}
