
\section{Discussion}\label{sec:discussion}

We described two essential techniques, viz.\ multi-shot and theory solving, for enhancing ASP solving
by different forms of hybridization.
While multi-shot solving allows for fine-grained control of ASP reasoning processes,
theory solving allows for refining basic ASP solving by incorporating foreign types of constraints.
%
Since ASP follows a model, ground, and solve methodology both techniques pervade the whole work-flow of ASP,
starting with extensions to the input language, over means for incremental and theory-enhanced grounding,
to stateful and theory-enhanced solving.
Multi-shot solving even adds a fourth dimension to control ASP reasoning processes.

Our focus on \clingo\ should not conceal other approaches to hybrid ASP solving.
Foremost,
\dlvhex~\cite{redl16a} builds upon \clingo's infrastructure to provide a higher level of hybridization via higher-order logic programs.
As well, \clingcon~\cite{bakaossc16a} and \lctocasp~\cite{cakaossc16a} rely on \clingo{} for extending ASP with linear constraints over integers.
Similar yet customized approaches include \adsolver~\cite{megezh08a}, \inca~\cite{drewal10a}, and \ezcsp~\cite{ballie17a}.
Another category of ASP systems,
such us \ezsmt~\cite{liesus16a}, \dingo~\cite{jalini11a}, and \aspmt~\cite{barlee14b}
translate ASP with constraints to SAT Modulo Theories (SMT~\cite{niolti06a}) and use appropriate back-ends.
Similarly, \mingo~\cite{lijani12a} translates to Mixed Integer Linear Programming (MILP)
and \aspartame~\cite{bageinospescsotawe15a} back to ASP using the order encoding~\cite{crabak94a,tatakiba09a}.

Theory propagators have recently also been added to the ASP solver \wasp~\cite{dorisc16a};
these can be made accessible via the theory language of Section~\ref{sec:language} along with the intermediate format described in
Appendix~\ref{sec:aspif}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: 
