
\paragraph{USA-Advisor}
\begin{itemize}
\item Description copied from \cite{balduccini10a}:
Problem instances from the task of planning for the Reaction Control System of the Space Shuttle \cite{balgelnog06a,nogueira03a}. 
A collection of problem instances from the domain of the RCS is publicy available, at 
http://www.krlab.cs.ttu.edu/Software/Download/,
together with the ASP encodings of the model of the RCS and of various reasoning modules.
The interesed reader may refer to \cite{nogueira03a} for a description of the instances.
\item 
For our evaluation, we used the public instances (2000) and for each 
one we consider plan lenghts 1 to 10. In total, 20000 runs. 
\item
First heuristic (Forward): Forward search on actions:
\begin{lstlisting}
_h(occurs(A,T),true,l-T) :- 
	 action_of(A,R), time(T).
\end{lstlisting}
\item
Second heuristic (Forward-Sub): The RCS is divided in three subsystems, 
fwd\_rcs, left\_rcs and right\_rcs, of which only the last two are interconnected.
With this heuristic we do forward search on actions subsystem by subsystem, in order. 
First we solve subsystem fwd\_rcs, then subsystem left\_rcs and finally subsystem right\_rcs:
\begin{lstlisting}
_heuristic(occurs(A,T),true,3*l-T) :- 
           action_of(A,R), time(T), R=fwd_rcs.
_heuristic(occurs(A,T),true,2*l-T) :- 
           action_of(A,R), time(T), R=left_rcs.
_heuristic(occurs(A,T),true,1*l-T) :- 
           action_of(A,R), time(T), R=right_rcs.
\end{lstlisting}
If the systems $R$ were represented by numbers 1, 2 and 3, we could simply write:
\begin{lstlisting}
_heuristic(occurs(A,T),true,R*l-T) :- 
           action_of(A,R), time(T).
\end{lstlisting}
\item All instances are solved very fast. 
I give for the 20000 runs the average for:
runsolver time, 
solving time that appears in clasp output, 
number of conflicts and number of choices. 
I also give the same data for satisfiable instances, 
and for unsatisfiable ones.
In Table \ref{tab:usa-advisor} i put a sample table with data of all instances  
and of satisfiable instances.
\item 
Follows the data so that a macro can be applied for formatting
\item
X,time,solve,conflicts,choices,time-sat,solve-sat,conflicts-sat,choices-sat,time-unsat,solve-unsat,conflicts-unsat,choices-unsat
clasp,0.2234,0.0087,40.8,1043.2,0.319,0.0215,98,2561.4,0.1632,0.0006,4.8,87.1
hclasp-forward,0.2241,0.0031,8.4,21.1,0.31,0.0017,5.7,21.8,0.17,0.004,10.1,20.6
hclasp-forward-sub,0.2223,0.0015,6.5,14.1,0.3096,0.0019,6.6,22.3,0.1673,0.0013,6.5,9
\item
Results for all instances:
Similar runsolver time, 
forward three times less solving-time than vsids, 
and forward-sub six times less solving-time than vsids. 
Forward much less conflicts and choices than vsids
(1 and 2 orders of magnitude, respectively), 
and forward-sub even improves a bit Forward.
\item
Results for satisfiable instances:
Similar runsolver time, 
forward and forward-sub one order of magnitude less solving-time than vsids.
For conflicts and choices the differences now are bigger, 
keeping 1 and 2 orders of magnitude.
\item
Results for unsatisfiable instances:
Similar runsolver time,
forward-sub double solving-time than vsids, 
and forward six times more solving-time than vsids.
For conflicts forward-sub does a bit more than vsids, 
and forward doubles vsids.
For choices forward-sub does the least (9) then forward does 20 and vsids moves to 90.
So vsids does more choices but less conflicts. 
forward and forward-sub one order of magnitude less solving-time than vsids.
For conflicts and choices the differences now are bigger, 
keeping 1 and 2 orders of magnitude.
\item Summary of the results: 
\begin{itemize}
\item
for runsolving time there are no differences because solving is extremely fast
\item 
for clasp-solving time on SAT instances hclasp in both cases wins by one order of magnitude,
while for UNSAT vsids is better, 2 times faster than forward-sub and six times faster than forward
\item 
hclasp does less conflicts on SAT instances (one order of magnitude) and does more conflicts on UNSAT instances
\item 
hclasp does always less choices, two orders of magnitude less for SAT instances, and one for UNSAT
\item forward is just a bit better than forward-sub on SAT, 
but forward-sub is better on UNSAT (three times less solving time, half choices and conflicts).
\end{itemize}
\item With tetris i got no improvements.

\end{itemize}

 \begin{table*}[ht]
   \centering
   \begin{tabular}{|c|cc|}
     \hline
system & SAT (time/solving time/conflicts/choices) & SAT (time/solving time/conflicts/choices) \\ \hline     
clasp & 0.2234/0.0087/40.8/1043.2 & 0.319/0.0215/98/2561.4/0.1632 \\ \hline
hclasp-forward & 0.2241/0.0031/8.4/21.1 & 0.31/0.0017/5.7/21.8/0.17 \\ \hline
hclasp-forward-subsystems & 0.2223/0.0015/6.5/14.1 & 0.3096/0.0019/6.6/22.3/0.1673 \\ \hline
   \end{tabular}
   \caption{}
   \label{tab:usa-advisor}
 \end{table*}
