\begin{thebibliography}{}

\bibitem[\protect\citeauthoryear{Bacchus and Kabanza}{2000}]{backab00a}
Bacchus, F., and Kabanza, F.
\newblock 2000.
\newblock Using temporal logics to express search control knowledge for
  planning.
\newblock {\em Artificial Intelligence} 116(1-2):123--191.

\bibitem[\protect\citeauthoryear{Balduccini}{2011}]{balduccini11b}
Balduccini, M.
\newblock 2011.
\newblock Learning and using domain-specific heuristics in {ASP} solvers.
\newblock {\em AI Communic.} 24(2):147--164.

\bibitem[\protect\citeauthoryear{Baral}{2003}]{baral02a}
Baral, C.
\newblock 2003.
\newblock {\em Knowledge Representation, Reasoning and Declarative Problem
  Solving}.
\newblock Cambridge University Press.

\bibitem[\protect\citeauthoryear{Biere \bgroup et al\mbox.\egroup
  }{2009}]{SATHandbook}
Biere, A.; Heule, M.; {van Maaren}, H.; and Walsh, T., eds.
\newblock 2009.
\newblock {\em Handbook of Satisfiability}, volume 185 of {\em Frontiers in
  Artificial Intelligence and Applications}.
\newblock IOS Press.

\bibitem[\protect\citeauthoryear{Calimeri \bgroup et al\mbox.\egroup
  }{2011}]{contest11a}
Calimeri, F. et al.
\newblock 2011.
\newblock The third answer set programming competition: Preliminary report of
  the system competition track.
\newblock In Delgrande and Faber \shortcite{lpnmr11},  388--403.

\bibitem[\protect\citeauthoryear{Castell \bgroup et al\mbox.\egroup
  }{1996}]{cacacale96a}
Castell, T.; Cayrol, C.; Cayrol, M.; and {Le Berre}, D.
\newblock 1996.
\newblock Using the {D}avis and {P}utnam procedure for an efficient computation
  of preferred models.
\newblock In Wahlster, W., ed., {\em Proceedings of the Twelfth European
  Conference on Artificial Intelligence (ECAI'96)},  350--354.
\newblock John Wiley \& sons.

\bibitem[\protect\citeauthoryear{Dechter}{2003}]{dechter03}
Dechter, R.
\newblock 2003.
\newblock {\em Constraint Processing}.
\newblock Morgan Kaufmann.

\bibitem[\protect\citeauthoryear{Delgrande and Faber}{2011}]{lpnmr11}
Delgrande, J., and Faber, W., eds.
\newblock 2011.
\newblock {\em Proceedings of the Eleventh International Conference on Logic
  Programming and Nonmonotonic Reasoning (LPNMR'11)}. Springer.

\bibitem[\protect\citeauthoryear{{Di Rosa}, Giunchiglia, and
  Maratea}{2010}]{rogima10a}
{Di Rosa}, E.; Giunchiglia, E.; and Maratea, M.
\newblock 2010.
\newblock Solving satisfiability problems with preferences.
\newblock {\em Constraints} 15(4):485--515.

\bibitem[\protect\citeauthoryear{Faber \bgroup et al\mbox.\egroup
  }{2007}]{falemari07a}
Faber, W.; Leone, N.; Maratea, M.; and Ricca, F.
\newblock 2007.
\newblock Experimenting with look-back heuristics for hard {ASP} programs.
\newblock In Baral, C.; Brewka, G.; and Schlipf, J., eds., {\em Proceedings of
  the Ninth International Conference on Logic Programming and Nonmonotonic
  Reasoning (LPNMR'07)},  110--122.
\newblock Springer.

\bibitem[\protect\citeauthoryear{Faber, Leone, and Pfeifer}{2001}]{falepf01a}
Faber, W.; Leone, N.; and Pfeifer, G.
\newblock 2001.
\newblock Experimenting with heuristics for answer set programming.
\newblock In Nebel, B., ed., {\em Proceedings of the Seventeenth International
  Joint Conference on Artificial Intelligence (IJCAI'01)},  635--640.
\newblock Morgan Kaufmann.

\bibitem[\protect\citeauthoryear{Freeman}{1995}]{freeman95a}
Freeman, J.
\newblock 1995.
\newblock {\em Improvements to Propositional Satisfiability Search Algorithms}.
\newblock Ph.D. Dissertation, University of Pennsylvania.

\bibitem[\protect\citeauthoryear{Gebser \bgroup et al\mbox.\egroup
  }{2010}]{geguivscsithve10a}
Gebser, M.; Guziolowski, C.; Ivanchev, M.; Schaub, T.; Siegel, A.; Thiele, S.;
  and Veber, P.
\newblock 2010.
\newblock Repair and prediction (under inconsistency) in large biological
  networks with answer set programming.
\newblock In Lin, F., and Sattler, U., eds., {\em Proceedings of the Twelfth
  International Conference on Principles of Knowledge Representation and
  Reasoning (KR'10)},  497--507.
\newblock AAAI Press.

\bibitem[\protect\citeauthoryear{Gebser \bgroup et al\mbox.\egroup
  }{2011}]{gekaknsc11a}
Gebser, M.; Kaminski, R.; Knecht, M.; and Schaub, T.
\newblock 2011.
\newblock plasp: A prototype for {PDDL}-based planning in {ASP}.
\newblock In Delgrande and Faber \shortcite{lpnmr11},  358--363.

\bibitem[\protect\citeauthoryear{Gebser \bgroup et al\mbox.\egroup
  }{2012}]{gekakasc12a}
Gebser, M.; Kaminski, R.; Kaufmann, B.; Schaub, T.
\newblock 2012.
\newblock {\em Answer Set Solving in Practice}.
\newblock  Morgan and Claypool.
% Synthesis Lectures on Artificial Intelligence and Machine Learning.


\bibitem[\protect\citeauthoryear{Giunchiglia and Maratea}{2012}]{giumar12a}
Giunchiglia, E., and Maratea, M.
\newblock 2012.
\newblock Algorithms for solving satisfiability problems with qualitative
  preferences.
\newblock In Erdem, E.; Lee, J.; Lierler, Y.; and Pearce, D., eds., {\em
  Correct Reasoning: Essays on Logic-Based {AI} in Honour of {V}ladimir
  {L}ifschitz},
  327--344.
\newblock Springer.

\bibitem[\protect\citeauthoryear{Goldberg and Novikov}{2002}]{golnov02a}
Goldberg, E., and Novikov, Y.
\newblock 2002.
\newblock {BerkMin}: A fast and robust {SAT} solver.
\newblock In {\em Proceedings of the Fifth Conference on Design, Automation and
  Test in Europe (DATE'02)},  142--149.
\newblock IEEE Computer Society Press.

\bibitem[\protect\citeauthoryear{hclasp\ignorespaces}{\ignorespaces}]{hclasp}
\newblock hclasp. \texttt{http://www.cs.uni-potsdam.de/hclasp}.

\bibitem[\protect\citeauthoryear{ICAPS\ignorespaces}{\ignorespaces}]{icaps-competition}
\newblock ICAPS. \texttt{http://ipc.icaps-conference.org}.

\bibitem[\protect\citeauthoryear{J{\"a}rvisalo, Junttila, and
  Niemel{\"a}}{2005}]{jajuni05a}
J{\"a}rvisalo, M.; Junttila, T.; and Niemel{\"a}, I.
\newblock 2005.
\newblock Unrestricted vs restricted cut in a tableau method for {B}oolean
  circuits.
\newblock {\em Annals of Mathematics and Artificial Intelligence}
  44(4):373--399.

\bibitem[\protect\citeauthoryear{Korf}{1985}]{korf85a}
Korf, R.
\newblock 1985.
\newblock Depth-first iterative-deepening: An optimal admissible tree search.
\newblock {\em Artificial Intelligence} 27(1):97--109.

\bibitem[\protect\citeauthoryear{Lifschitz}{2002}]{lifschitz02a}
Lifschitz, V.
\newblock 2002.
\newblock Answer set programming and plan generation.
\newblock {\em Artificial Intelligence} 138(1-2):39--54.

\bibitem[\protect\citeauthoryear{Marques-Silva and Sakallah}{1999}]{marsak99a}
Marques-Silva, J., and Sakallah, K.
\newblock 1999.
\newblock {GRASP}: A search algorithm for propositional satisfiability.
\newblock {\em IEEE Transactions on Computers} 48(5):506--521.

\bibitem[\protect\citeauthoryear{Moskewicz \bgroup et al\mbox.\egroup
  }{2001}]{momazhzhma01a}
Moskewicz, M.; Madigan, C.; Zhao, Y.; Zhang, L.; and Malik, S.
\newblock 2001.
\newblock Chaff: Engineering an efficient {SAT} solver.
\newblock In {\em Proceedings of the Thirty-eighth Conference on Design
  Automation (DAC'01)},  530--535.
\newblock ACM Press.

\bibitem[\protect\citeauthoryear{Pipatsrisawat and Darwiche}{2007}]{pipdar07a}
Pipatsrisawat, K., and Darwiche, A.
\newblock 2007.
\newblock A lightweight component caching scheme for satisfiability solvers.
\newblock In Marques-Silva, J., and Sakallah, K., eds., {\em Proceedings of the
  Tenth International Conference on Theory and Applications of Satisfiability
  Testing (SAT'07)},
  294--299.
\newblock Springer.

\bibitem[\protect\citeauthoryear{Pretolani}{1996}]{pretolani96a}
Pretolani, D.
\newblock 1996.
\newblock Efficiency and stability of hypergraph {SAT} algorithms.
\newblock In Johnson, D., and Trick, M., eds., {\em DIMACS Series in Discrete
  Mathematics and Theoretical Computer Science}, volume~26,  479--498.
\newblock American Mathematical Society.

\bibitem[\protect\citeauthoryear{Rintanen}{2011}]{rintanen11a}
Rintanen, J.
\newblock 2011.
\newblock Planning with {SAT}, admissible heuristics and {A}$^*$.
\newblock In Walsh \shortcite{ijcai11},  2015--2020.

\bibitem[\protect\citeauthoryear{Rintanen}{2012}]{rintanen12a}
Rintanen, J.
\newblock 2012.
\newblock Planning as satisfiability: heuristics.
\newblock {\em Artificial Intelligence} 193:45--86.

\bibitem[\protect\citeauthoryear{Schaub and Thiele}{2009}]{schthi09a}
Schaub, T., and Thiele, S.
\newblock 2009.
\newblock Metabolic network expansion with {ASP}.
\newblock In Hill, P., and Warren, D., eds., {\em Proceedings of the
  Twenty-fifth International Conference on Logic Programming (ICLP'09)},  312--326.
\newblock Springer.

\bibitem[\protect\citeauthoryear{Siddiqi}{2011}]{sidiqqi11a}
Siddiqi, S.
\newblock 2011.
\newblock Computing minimum-cardinality diagnoses by model relaxation.
\newblock In Walsh \shortcite{ijcai11},  1087--1092.

\bibitem[\protect\citeauthoryear{Sierra-Santib{\'a}{\~n}ez}{2004}]{sierra04a}
Sierra-Santib{\'a}{\~n}ez, J.
\newblock 2004.
\newblock Heuristic planning: A declarative approach based on strategies for
  action selection.
\newblock {\em Artificial Intelligence} 153(1-2):307--337.

\bibitem[\protect\citeauthoryear{Walsh}{2011}]{ijcai11}
Walsh, T., ed.
\newblock 2011.
\newblock {\em Proceedings of the Twenty-second International Joint Conference
  on Artificial Intelligence (IJCAI'11)}. IJCAI/AAAI.

\bibitem[\protect\citeauthoryear{Zhang \bgroup et al\mbox.\egroup
  }{2001}]{zamamoma01a}
Zhang, L.; Madigan, C.; Moskewicz, M.; and Malik, S.
\newblock 2001.
\newblock Efficient conflict driven learning in a {B}oolean satisfiability
  solver.
\newblock In {\em Proceedings of the International Conference on Computer-Aided
  Design (ICCAD'01)},  279--285.
\newblock ACM Press.

\end{thebibliography}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: 
