% -----------------------------------------------------------------------------------------------------------------------
% -----------------------------------------------------------------------------------------------------------------------
% -----------------------------------------------------------------------------------------------------------------------

%\begin{frame}[fragile]{
\begin{block}{Heuristics in ASP for PDDL planning}
%}

\dosubtitle{Simple STRIPS planner}

\begin{columns} \begin{column}{.02\linewidth}\end{column}
\begin{column}{.98\linewidth}


\begin{semiverbatim}
time(1..last).    

holds(P,0) :- init(P).

1 \{ occurs(A,T) : action(A) \} 1 :- time(T).
:-  occurs(A,T), pre(A,F), not holds(F,T-1).

holds(F,T)  :- holds(F,T-1), not nholds(F,T), time(T).
holds(F,T)  :- occurs(A,T), add(A,F).
nholds(F,T) :- occurs(A,T), del(A,F).

:- query(F), not holds(F,last).

\end{semiverbatim}

\end{column} \end{columns}


\dosubtitle{Heuristic rules}
\begin{columns} \begin{column}{.02\linewidth}\end{column}
\begin{column}{.98\linewidth}

Add \verb|5| to the initial heuristic value of actions

\verb|_heuristic(occurs(A,T),init,5) :- action(A), time(T).|

\bigskip \bigskip

Multiply the heuristic value of actions at \verb|T| by \verb|T|

\verb|_heuristic(occurs(A,T),factor,T) :- action(A), time(T).|

\bigskip \bigskip

Decide first on actions

\verb|_heuristic(occurs(A,T),level,1) :- action(A), time(T).|

\bigskip \bigskip

Do a forward search on actions

\verb|_heuristic(occurs(A,T),level,last-T+1) :- action(A), time(T).|

\bigskip \bigskip

If deciding on actions, assign them to false

\verb|_heuristic(occurs(A,T),sign,-1) :- action(A), time(T).|


\end{column} \end{columns}

\fourskip 

\dosubtitle{New modifiers definition}
\begin{columns} \begin{column}{.02\linewidth}\end{column}
\begin{column}{.98\linewidth}

Modifier \verb|true| gives a rank and a positive sign, 
\verb|false| gives a rank and a negative sign

\verb|_heuristic(X,level,Y) :- _heuristic(X,true,Y).|
\verb|_heuristic(X,sign,1)  :- _heuristic(X,true,Y).|
\verb|_heuristic(X,level,Y) :- _heuristic(X,false,Y).|
\verb|_heuristic(X,sign,-1) :- _heuristic(X,false,Y).|

\bigskip \bigskip

Do a forward search on true actions

\verb|_heuristic(occurs(A,T),true,last-T+1) :- action(A), time(T).|



%_heuristic(occurs(A,T),factor,2) :- action(A), time(T).
%_heuristic(occurs(A,T),factor,2) :- action(A), time(T).
%\end{semiverbatim}

%%}\only<5>{_heuristic(A,level,V) :- _heuristic(A,true, V).
%%_heuristic(A,sign, 1) :- _heuristic(A,true, V).
%%}\only<6>{_heuristic(A,level,V) :- _heuristic(A,false,V).
%%_heuristic(A,sign,-1) :- _heuristic(A,false,V).
%%}\only<7>{_heuristic(holds(F,T-1),true, t-T+1) :- holds(F,T).
%%_heuristic(holds(F,T-1),false,t-T+1) :-
%%                fluent(F), time(T), not holds(F,T).}

\end{column} \end{columns}

%%\end{frame}

\end{block}



% -----------------------------------------------------------------------------------------------------------------------
% -----------------------------------------------------------------------------------------------------------------------
% -----------------------------------------------------------------------------------------------------------------------

