\begin{thebibliography}{}

\bibitem[\protect\citeauthoryear{Banbara, Gebser, Inoue, Ostrowski, Peano,
  Schaub, Soh, Tamura, and Weise}{Banbara
  et~al\mbox{.}}{2015}]{bageinospescsotawe15a}
{\sc Banbara, M.}, {\sc Gebser, M.}, {\sc Inoue, K.}, {\sc Ostrowski, M.}, {\sc
  Peano, A.}, {\sc Schaub, T.}, {\sc Soh, T.}, {\sc Tamura, N.}, {\sc and} {\sc
  Weise, M.} 2015.
\newblock aspartame: Solving constraint satisfaction problems with answer set
  programming.
\newblock In {\em Proceedings of the Thirteenth International Conference on
  Logic Programming and Nonmonotonic Reasoning (LPNMR'15)}, {F.~Calimeri},
  {G.~Ianni}, {and} {M.~Truszczy{\'n}ski}, Eds. Lecture Notes in Artificial
  Intelligence, vol. 9345. Springer-Verlag, 112--126.

\bibitem[\protect\citeauthoryear{Banbara, Kaufmann, Ostrowski, and
  Schaub}{Banbara et~al\mbox{.}}{2017}]{bakaossc16a}
{\sc Banbara, M.}, {\sc Kaufmann, B.}, {\sc Ostrowski, M.}, {\sc and} {\sc
  Schaub, T.} 2017.
\newblock Clingcon: The next generation.
\newblock {\em Theory and Practice of Logic Programming\/}.
\newblock To appear.

\bibitem[\protect\citeauthoryear{Baral}{Baral}{2003}]{baral02a}
{\sc Baral, C.} 2003.
\newblock {\em Knowledge Representation, Reasoning and Declarative Problem
  Solving}.
\newblock Cambridge University Press.

\bibitem[\protect\citeauthoryear{Barrett, Sebastiani, Seshia, and
  Tinelli}{Barrett et~al\mbox{.}}{2009}]{baseseti09a}
{\sc Barrett, C.}, {\sc Sebastiani, R.}, {\sc Seshia, S.}, {\sc and} {\sc
  Tinelli, C.} 2009.
\newblock Satisfiability modulo theories.
\newblock In {\em Handbook of Satisfiability}, {A.~Biere}, {M.~Heule}, {H.~{van
  Maaren}}, {and} {T.~Walsh}, Eds. Frontiers in Artificial Intelligence and
  Applications, vol. 185. IOS Press, Chapter~26, 825--885.

\bibitem[\protect\citeauthoryear{Bartholomew and Lee}{Bartholomew and
  Lee}{2014}]{barlee14b}
{\sc Bartholomew, M.} {\sc and} {\sc Lee, J.} 2014.
\newblock System aspmt2smt: Computing {ASPMT} theories by {SMT} solvers.
\newblock In {\em Proceedings of the Fourteenth European Conference on Logics
  in Artificial Intelligence (JELIA'14)}, {E.~Ferm\'{e}} {and} {J.~Leite}, Eds.
  Lecture Notes in Artificial Intelligence, vol. 8761. Springer-Verlag,
  529--542.

\bibitem[\protect\citeauthoryear{Cabalar, Otero, and Pose}{Cabalar
  et~al\mbox{.}}{2000}]{caotpo00a}
{\sc Cabalar, P.}, {\sc Otero, R.}, {\sc and} {\sc Pose, S.} 2000.
\newblock Temporal constraint networks in action.
\newblock In {\em Proceedings of the Fourteenth European Conference on
  Artificial Intelligence (ECAI'00)}, {W.~Horn}, Ed. IOS Press, 543--547.

\bibitem[\protect\citeauthoryear{Carro and King}{Carro and
  King}{2016}]{iclp-lipics16}
{\sc Carro, M.} {\sc and} {\sc King, A.}, Eds. 2016.
\newblock {\em Technical Communications of the Thirty-second International
  Conference on Logic Programming (ICLP'16)}. Vol.~52. Open Access Series in
  Informatics (OASIcs).

\bibitem[\protect\citeauthoryear{Cotton and Maler}{Cotton and
  Maler}{2006}]{cotmal06a}
{\sc Cotton, S.} {\sc and} {\sc Maler, O.} 2006.
\newblock Fast and flexible difference constraint propagation for {DPLL (T)}.
\newblock In {\em Proceedings of the Ninth International Conference on Theory
  and Applications of Satisfiability Testing (SAT'06)}, {A.~Biere} {and}
  {C.~Gomes}, Eds. Lecture Notes in Computer Science, vol. 4121.
  Springer-Verlag, 170--183.

\bibitem[\protect\citeauthoryear{Crawford and Baker}{Crawford and
  Baker}{1994}]{crabak94a}
{\sc Crawford, J.} {\sc and} {\sc Baker, A.} 1994.
\newblock Experimental results on the application of satisfiability algorithms
  to scheduling problems.
\newblock In {\em Proceedings of the Twelfth National Conference on Artificial
  Intelligence (AAAI'94)}, {B.~Hayes-Roth} {and} {R.~Korf}, Eds. AAAI Press,
  1092--1097.

\bibitem[\protect\citeauthoryear{Dantzig}{Dantzig}{1963}]{dantzig63a}
{\sc Dantzig, G.} 1963.
\newblock {\em Linear Programming and Extensions}.
\newblock Princeton University Press.

\bibitem[\protect\citeauthoryear{{De Rosis}, Eiter, Redl, and Ricca}{{De Rosis}
  et~al\mbox{.}}{}]{roeireri15a}
{\sc {De Rosis}, A.}, {\sc Eiter, T.}, {\sc Redl, C.}, {\sc and} {\sc Ricca,
  F.}
\newblock Constraint answer set programming based on {HEX}-programs.

\bibitem[\protect\citeauthoryear{Drescher and Walsh}{Drescher and
  Walsh}{2010}]{drewal10a}
{\sc Drescher, C.} {\sc and} {\sc Walsh, T.} 2010.
\newblock A translational approach to constraint answer set solving.
\newblock {\em Theory and Practice of Logic Programming\/}~{\em 10,\/}~4-6,
  465--480.

\bibitem[\protect\citeauthoryear{Gebser, Kaminski, Kaufmann, Ostrowski, Schaub,
  and Wanko}{Gebser et~al\mbox{.}}{2016}]{gekakaosscwa16a}
{\sc Gebser, M.}, {\sc Kaminski, R.}, {\sc Kaufmann, B.}, {\sc Ostrowski, M.},
  {\sc Schaub, T.}, {\sc and} {\sc Wanko, P.} 2016.
\newblock Theory solving made easy with clingo~5.
\newblock See \citeN{iclp-lipics16}, 2:1--2:15.

\bibitem[\protect\citeauthoryear{Gebser, Kaminski, Kaufmann, and Schaub}{Gebser
  et~al\mbox{.}}{2014}]{gekakasc14b}
{\sc Gebser, M.}, {\sc Kaminski, R.}, {\sc Kaufmann, B.}, {\sc and} {\sc
  Schaub, T.} 2014.
\newblock \textit{Clingo} = {ASP} + control: Preliminary report.
\newblock In {\em Technical Communications of the Thirtieth International
  Conference on Logic Programming (ICLP'14)}, {M.~Leuschel} {and}
  {T.~Schrijvers}, Eds. Theory and Practice of Logic Programming, Online
  Supplement, vol. arXiv:1405.3694v1.
\newblock Available at \url{http://arxiv.org/abs/1405.3694v1}.

\bibitem[\protect\citeauthoryear{Gebser, Kaufmann, and Schaub}{Gebser
  et~al\mbox{.}}{2012}]{gekasc09c}
{\sc Gebser, M.}, {\sc Kaufmann, B.}, {\sc and} {\sc Schaub, T.} 2012.
\newblock Conflict-driven answer set solving: From theory to practice.
\newblock {\em Artificial Intelligence\/}~{\em 187-188}, 52--89.

\bibitem[\protect\citeauthoryear{Gelfond and Lifschitz}{Gelfond and
  Lifschitz}{1991}]{gellif91a}
{\sc Gelfond, M.} {\sc and} {\sc Lifschitz, V.} 1991.
\newblock Classical negation in logic programs and disjunctive databases.
\newblock {\em New Generation Computing\/}~{\em 9}, 365--385.

\bibitem[\protect\citeauthoryear{Goldberg}{Goldberg}{1991}]{goldberg91a}
{\sc Goldberg, D.} 1991.
\newblock What every computer scientist should know about floating-point
  arithmetic.
\newblock {\em ACM Computing Surveys (CSUR)\/}~{\em 23,\/}~1, 5--48.

\bibitem[\protect\citeauthoryear{Janhunen, Liu, and Niemelä}{Janhunen
  et~al\mbox{.}}{2011}]{jalini11a}
{\sc Janhunen, T.}, {\sc Liu, G.}, {\sc and} {\sc Niemelä, I.} 2011.
\newblock Tight integration of non-ground answer set programming and
  satisfiability modulo theories.
\newblock In {\em Proceedings of the First Workshop on Grounding and
  Transformation for Theories with Variables (GTTV'11)}, {P.~Cabalar},
  {D.~Mitchell}, {D.~Pearce}, {and} {E.~Ternovska}, Eds. 1--13.

\bibitem[\protect\citeauthoryear{Lierler and Susman}{Lierler and
  Susman}{2016}]{liesus16a}
{\sc Lierler, Y.} {\sc and} {\sc Susman, B.} 2016.
\newblock {SMT}-based constraint answer set solver {EZSMT} (system
  description).
\newblock See \citeN{iclp-lipics16}, 1:1--1:15.

\bibitem[\protect\citeauthoryear{Liu, Janhunen, and Niemelä}{Liu
  et~al\mbox{.}}{2012}]{lijani12a}
{\sc Liu, G.}, {\sc Janhunen, T.}, {\sc and} {\sc Niemelä, I.} 2012.
\newblock Answer set programming via mixed integer programming.
\newblock In {\em Proceedings of the Thirteenth International Conference on
  Principles of Knowledge Representation and Reasoning (KR'12)}, {G.~Brewka},
  {T.~Eiter}, {and} {S.~McIlraith}, Eds. AAAI Press, 32--42.

\bibitem[\protect\citeauthoryear{Simons, Niemelä, and Soininen}{Simons
  et~al\mbox{.}}{2002}]{siniso02a}
{\sc Simons, P.}, {\sc Niemelä, I.}, {\sc and} {\sc Soininen, T.} 2002.
\newblock Extending and implementing the stable model semantics.
\newblock {\em Artificial Intelligence\/}~{\em 138,\/}~1-2, 181--234.

\bibitem[\protect\citeauthoryear{Soh, Inoue, Tamura, Banbara, and
  Nabeshima}{Soh et~al\mbox{.}}{2010}]{sointabana10a}
{\sc Soh, T.}, {\sc Inoue, K.}, {\sc Tamura, N.}, {\sc Banbara, M.}, {\sc and}
  {\sc Nabeshima, H.} 2010.
\newblock A {SAT}-based method for solving the two-dimensional strip packing
  problem.
\newblock {\em Fundamenta Informaticae\/}~{\em 102,\/}~3-4, 467--487.

\bibitem[\protect\citeauthoryear{Taillard}{Taillard}{1993}]{taillard93a}
{\sc Taillard, E.} 1993.
\newblock Benchmarks for basic scheduling problems.
\newblock {\em European Journal of Operational Research\/}~{\em 64,\/}~2,
  278--285.

\bibitem[\protect\citeauthoryear{van Loon}{van Loon}{1981}]{loon81}
{\sc van Loon, J.} 1981.
\newblock Irreducibly inconsistent systems of linear inequalities.
\newblock In {\em European Journal of Operational Research}. Vol.~3. Elsevier
  Science, 283--288.

\end{thebibliography}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: 
